---
title: "Chapter 4: Building Causal Diagrams fFrom Scratch"
output: html_notebook
---

This is the [R Markdown](http://rmarkdown.rstudio.com) Notebook for chapter 4. This script loads the example data, based on data from Antonio, de Almeida & Nunes,  "Hotel booking demand datasets", Data in Brief, 2019.  https://doi.org/10.1016/j.dib.2018.11.126. and analyzes it.

# Business problem and data setup
## Data and libraries
```{r}
# Common libraries
suppressMessages(suppressWarnings(library(tidyverse)))
library(boot) #Required for Bootstrap simulations
library(rstudioapi) #To load data from local folder
library(ggpubr) #To generate multi-plots

# Chapter-specific libraries
library(rcompanion) # For Cramer V correlation coefficient
library(corrplot) # For Pearson's correlation coefficient, cor()
#library(car) #For vif diagnostic function

### Setting the working directory to the parent folder of this script (Rstudio only)
sourceDir <- rstudioapi::getActiveDocumentContext()$path %>% str_extract("^.+/")
setwd(sourceDir)

set.seed(1234)
options(scipen=10)
```

```{r}
#Loading the data
dat <- read_csv("chap4-hotel_booking_case_study.csv", 
                col_types = cols(
                  NRDeposit = col_factor(),
                  IsCanceled = col_factor(),
                  DistributionChannel = col_factor(),
                  MarketSegment = col_factor(),
                  CustomerType = col_factor(),
                  Children = col_double(),
                  ADR = col_double(),
                  PreviousCancellations = col_factor(),
                  IsRepeatedGuest = col_factor(),
                  Country = col_character(),
                  Quarter = col_factor(),
                  Year = col_double()))

#Reducing the number of values for Country of origin by keeping most frequent countries only
#and aggregating the remaining ones under "Other"
countries <- dat %>% 
  group_by(Country) %>% 
  summarize(pct = n()/nrow(dat))
topcountries <- countries %>% 
  filter(pct >= 0.01)
dat <- dat %>%
  mutate(Country = ifelse(Country %in% topcountries$Country, Country, "Other")) %>%
  mutate(Country = as.factor(Country))
```
## Understanding the relationship of interest
```{r}
#Cancellation rate by deposit types
with(dat, table(NRDeposit, IsCanceled))
```


```{r}
with(dat, prop.table(table(NRDeposit, IsCanceled), 1))
```


```{r}
with(dat, rcompanion::cramerV(NRDeposit, IsCanceled))
```
# Validate observable variables to include based on data
## Relationships between numeric variables
```{r}
dat <- dat %>%
  rename(CustTyp= CustomerType) %>%
  rename(DistCh = DistributionChannel) %>%
  rename(RepGst = IsRepeatedGuest) %>%
  rename(MktSgmt = MarketSegment) %>%
  rename(IsCanc = IsCanceled) %>%
  rename(PrevCan = PreviousCancellations) %>%
  rename(NRDep = NRDeposit)
```

```{r}
num_dat <- dat %>%
  mutate(NRDep = as.integer(NRDep)) %>%
  mutate(IsCanc = as.integer(IsCanc)) %>%
  mutate(PrevCan = as.integer(PrevCan)) %>%
  mutate(RepGst = as.integer(RepGst)) %>%
  select_if(function(x) is.numeric(x)|is.integer(x))
```

```{r}
num_cor <- cor(num_dat)
num_cor
```
```{r}
corrplot.mixed(num_cor, lower.col = "black", number.cex = .7)
```
## Relationships between categorical variables
```{r}
### Calculating correlation table for all categorical variables

cat_corr_fun <- function(dat){
  # Renaming variables to shorten them
  
  #Going through all categorical variables
  corr_list <- list()
  n <- 1
  for(varI in colnames(dat)){
    for(varJ in colnames(dat)){
      #Factor to factor correlation
      if(is.factor(dat[[varI]]) & is.factor(dat[[varJ]]) & varI != varJ){
        corr = cramerV(dat[[varI]], dat[[varJ]])
        corr_list[[n]] <- data.frame(
          varI = as.character(varI),
          varJ = as.character(varJ),
          corr = corr)
        n <- n + 1
      }
    }
  }
  corr_df = bind_rows(corr_list) %>% 
    spread(varJ, corr)
  
  corr_df <- corr_df %>%
    relocate(NRDep, .after = varI) %>%
    mutate(varI = as.character(varI))
  
  cat_corr_mat <- as.matrix(corr_df %>% select(-varI))
  row.names(cat_corr_mat) <- corr_df$varI
  
  return(cat_corr_mat)
}

cat_corr_mat <- cat_corr_fun(dat)

corrplot.mixed(cat_corr_mat, lower.col = "black", number.cex = .7)
```
```{r}
##### Correlations between numeric and categorical variables #####
dat %>% group_by(CustTyp) %>% summarize(ADR = mean(ADR))
```


```{r}
dat %>% group_by(Country) %>% summarize(Year = mean(Year))
```

```{r}
ggplot(data=dat, aes(x=CustTyp, y = Year)) +geom_boxplot()
```

```{r}
summary(lm(Year~CustTyp, data=dat))
```

```{r}
# Regression analysis #####

#Single explanatory variable
mod1 <- glm(IsCanc ~ NRDep, 
            data = dat, family = binomial(link = "logit"))
summary(mod1)

```

```{r}
#Complete model
mod_all <- glm(IsCanc ~ ., 
                   data = dat, family = binomial(link = "logit"))
summary(mod_all)
```
