---
title: "Solutions for chapter 11 exercises"
output: html_notebook
---

# Set up

```{r}
options(scipen=10)

library(tidyverse)
library(rstudioapi)
library(boot)

### Setting the working directory to the parent folder of this script (Rstudio only)
sourceDir <- rstudioapi::getActiveDocumentContext()$path %>% str_extract("^.+/")
setwd(sourceDir)
```

# Reading the data

```{r}
data <- read_csv("Bertrand_Mullainathan_exercises_data.csv")
```

```{r}
#Reformating variables as factor
data <- data %>%
  mutate(race = factor(race, levels = c('w','b'))) %>%
  mutate(sex = factor(sex, levels = c('m','f'))) %>%
  mutate(college = factor(college, levels = c(0,1))) %>%
  mutate(eoe = factor(eoe, levels = c(0,1))) %>%
  mutate(wanted = factor(wanted))

# Filtering to secretary positions only
secretary_data <- data %>%
  filter(wanted=='secretary')
```

# 1. Visualizing the moderation. 

Replicate figure 11-3 with the EOE status as the x-variable, the call rate as the y-variable and race as the group variable. Interpret the figure.

Going from EOE=0 to EOE=1, the lines cross, meaning that the gap between sexes is reverted. Non-EOE firms call back male candidates more than female candidates whereas EOE firms call back female candidates more than male candidates.  

```{r}
viz_fun <- function(dat){
  #Representing visually moderation
  summary_data <- dat %>%
    group_by(sex, eoe) %>%
    summarize(cnt = n(),
              call_rate = mean(call)) %>%
    data.frame() 
  
  mod_p0 <- ggplot(summary_data, aes(x=eoe, y=call_rate, 
                                     group=sex)) + 
    geom_point() + geom_line(aes(lty=sex)) +
    scale_linetype_manual(values=c("dotted", "solid")) + theme_classic()
  mod_p0
}
viz_fun(secretary_data)
```
# 2 Measuring the moderation

2. a. Run a regression of call on sex and EOE status, with a moderation term between them. 

The moderation term is large and positive, suggesting that the EOE status increases the call back rate for female candidates. 

```{r}
mod <- glm(call~sex*eoe, data=secretary_data, family = 'binomial')
summary(mod)
```

2. b. Create the metric function for the regression in a) and then calculate the corresponding 80%-CI.

The 80%-CI is approximately [12.7; 15.1]. This is far from zero, so we can be confident that the effect is robust at the 80% level.  

```{r}
#Metric function

log_reg_fun <- function(dat){
  #Running logistic regression
  mod <- glm(call~sex*eoe, dat, family = 'binomial')
  summ <- summary(mod)
  metric <- summ$coefficients['sexf:eoe1', 'Estimate']
  return(metric)
}

#Bootstrap CI function
boot_CI_fun <- function(dat, metric_fun, B = 1000, conf.level = 0.8){
  #Setting the number of bootstrap samples
  boot_metric_fun <- function(dat, J){
    boot_dat <- dat[J,]
    return(metric_fun(boot_dat))
  }
  boot.out <- boot(data=dat, statistic=boot_metric_fun, R=B)
  confint <- boot.ci(boot.out, conf = conf.level, type = c('perc'))
  CI <- confint$percent[c(4,5)]
  return(CI)
}
set.seed(1)
boot_CI <- boot_CI_fun(dat=secretary_data, metric_fun=log_reg_fun)
```