---
title: "Solutions for chapter 12 exercises"
output: html_notebook
---

# Set up

```{r}
options(scipen=10)

library(tidyverse)
library(rstudioapi)
library(boot)

### Setting the working directory to the parent folder of this script (Rstudio only)
sourceDir <- rstudioapi::getActiveDocumentContext()$path %>% str_extract("^.+/")
setwd(sourceDir)

```

# Reading the data

```{r}
data <- read_csv("Karlan_List_exercises_data.csv")
```

```{r}
#Reframing variables as factor
data <- data %>%
  mutate(group = factor(group, levels=c('ctrl', 'treat1', 'treat2', 'treat3'))) %>%
  mutate(gender = factor(gender, levels=c('male', 'female', 'couple'))) %>%
  mutate(state_pol = factor(state_pol, levels=c('red','blue'))) %>%
  mutate(county_pol = factor(county_pol, levels=c('red', 'blue')))
summary(data)
```

# Setting up Bootstrap CI function

```{r}
#Bootstrap CI function
boot_CI_fun <- function(dat, metric_fun, B = 100, conf.level = 0.9){
  #Setting the number of bootstrap samples
  boot_metric_fun <- function(dat, J){
    boot_dat <- dat[J,]
    return(metric_fun(boot_dat))
  }
  boot.out <- boot(data=dat, statistic=boot_metric_fun, R=B)
  confint <- boot.ci(boot.out, conf = conf.level, type = c('perc'))
  CI <- confint$percent[c(4,5)]
  return(CI)
}
```

1. Calculate the total effect of gender == female on the amount given by running a regression including other covariates but excluding the tentative mediator.

The 90%-CI for the total effect of female donors is approx. [-0.25; 0.02].

```{r}
# Function definitions

#Metric functions
lin_reg_fun <- function(dat){
  #Running logistic regression
  mod <- lm(amount~group+gender+state_pol+county_pol, dat)
  summ <- summary(mod)
  metric <- summ$coefficients['genderfemale', 'Estimate']
  return(metric)
}

set.seed(1)
boot_CI_f_tot <- boot_CI_fun(dat=data, metric_fun=lin_reg_fun, conf.level = 0.9, B = 1e3)
```

2. Calculate the direct effect of gender == female on the amount given by running a regression including other covariates and the tentative mediator.

The 90%-CI for the direct effect of female donors is approx. [-0.2869;-0.0259].

```{r}
#Metric functions
lin_reg_fun <- function(dat){
  #Running logistic regression
  mod <- lm(amount~group+gender+state_pol+county_pol+freq+dormant, dat)
  summ <- summary(mod)
  metric <- summ$coefficients['genderfemale', 'Estimate']
  return(metric)
}

set.seed(1)
boot_CI_f_dir <- boot_CI_fun(dat=data, metric_fun=lin_reg_fun, conf.level = 0.9, B = 1e3)
```

3.	Calculate the indirect (mediated) effect of gender == female on the amount given by running the regression of the mediator on gender == female and a regression of the final effect on the mediator, the original cause of interest and relevant covariates, then multiplying the relevant coefficients (warning: do this multiplication inside the Bootstrap loop, not outside of it) 

The 90%-CI for the mediated effect of female donors is approx. [0.0207;0.0465].

```{r}
lin_reg_fun_med <- function(dat){
  mod1 <- lm(freq~gender+state_pol+county_pol, dat)
  summ1 <- summary(mod1)
  metric1 <- summ1$coefficients['genderfemale', 'Estimate']
  
  mod2 <- lm(amount~group+gender+state_pol+county_pol+freq+dormant, dat)
  summ2 <- summary(mod2)
  metric2 <- summ2$coefficients['freq', 'Estimate']
  
  return(metric1*metric2)
}

set.seed(1)
boot_CI_f_med <- boot_CI_fun(dat=data, metric_fun=lin_reg_fun_med, conf.level = 0.9, B = 1e3)
```

4.	Calculate the percentage mediated by running the same regressions as in 3 but also the regression for the total effect and determining the percentage mediated in each loop. What is your conclusion?

The 90%-CI for the percentage mediated effect of female donors is approx. [-164%; 85%], with a central value of -30%. The direct and indirect effect are going in opposite directions and mostly cancel each other (a complete cancellation would be a percentage mediated of -100%)

```{r}
lin_reg_fun_perc_med <- function(dat){
  mod1 <- lm(freq~gender+state_pol+county_pol, dat)
  summ1 <- summary(mod1)
  metric1 <- summ1$coefficients['genderfemale', 'Estimate']
  
  mod2 <- lm(amount~group+gender+state_pol+county_pol+freq+dormant, dat)
  summ2 <- summary(mod2)
  metric2 <- summ2$coefficients['freq', 'Estimate']
  
  mod <- lm(amount~group+gender+state_pol+county_pol, dat)
  summ <- summary(mod)
  metric <- summ$coefficients['genderfemale', 'Estimate']
  
  return(metric1*metric2/metric)
}

set.seed(1)
mean_perc_med <- lin_reg_fun_perc_med(data)
boot_CI_f_perc_med <- boot_CI_fun(dat=data, metric_fun=lin_reg_fun_perc_med, conf.level = 0.9, B = 1e3)
```








