---
title: "Solutions for chapter 11 exercises"
output: html_notebook
---

# Set up

```{r}
options(scipen=10)

library(tidyverse)
library(rstudioapi)
library(boot)

### Setting the working directory to the parent folder of this script (Rstudio only)
sourceDir <- rstudioapi::getActiveDocumentContext()$path %>% str_extract("^.+/")
setwd(sourceDir)
```

# Reading the data

```{r}
data <- read_csv("Bertrand_Mullainathan_exercises_data.csv")
```

```{r}
#Reformating variables as factor
data <- data %>%
  mutate(race = factor(race, levels = c('w','b'))) %>%
  mutate(gender = factor(gender, levels = c('m','f'))) %>%
  mutate(college = factor(college, levels = c(0,1))) %>%
  mutate(eoe = factor(eoe, levels = c(0,1))) %>%
  mutate(wanted = factor(wanted))
```

# 1. Visualizing the moderation. 

Replicate figure 11-3 with the EOE status as the x-variable, the call rate as the y-variable and gender as the group variable. Interpret the figure.

Going from EOE=0 to EOE=1, the lines cross, meaning that the gap between genders is reverted. Non-EOE firms call back female candidates more than male candidates whereas EOE firms call back male candidates more than female candidates. But the lines are pretty close to each other. The CI will tell us if that result is robust.

```{r}
viz_fun <- function(dat){
  #Representing visually moderation
  summary_data <- dat %>%
    group_by(gender, eoe) %>%
    summarize(cnt = n(),
              call_rate = mean(call)) %>%
    data.frame() 
  
  mod_p0 <- ggplot(summary_data, aes(x=eoe, y=call_rate, 
                                     group=gender)) + 
    geom_point() + geom_line(aes(lty=gender)) +
    scale_linetype_manual(values=c("dotted", "solid")) + theme_classic() +
    ylim(c(0,1))
  mod_p0
}
viz_fun(data)
```
# 2. Measuring the moderation

a. Run a regression of call on gender and EOE status, with a moderation term between them. 

The moderation term is negative, confirming that the EOE status decreases the relative call back rate for female candidates compared to male. 

```{r}
mod <- glm(call~gender*eoe, data=data, family = 'binomial')
summary(mod)
```

2. b. Create the metric function for the regression in a) and then calculate the corresponding 90%-CI.

The 90%-CI is approximately [-0.93; -0.05]. The CI doesn't cross zero, even with a large number of Bootstrap samples, so the effect is robust at the 90% level.  

```{r}
#Metric function

log_reg_fun <- function(dat){
  #Running logistic regression
  mod <- glm(call~gender*eoe, dat, family = 'binomial')
  summ <- summary(mod)
  metric <- summ$coefficients['genderf:eoe1', 'Estimate']
  return(metric)
}

#Bootstrap CI function
boot_CI_fun <- function(dat, metric_fun, B = 1000, conf.level = 0.9){
  #Setting the number of bootstrap samples
  boot_metric_fun <- function(dat, J){
    boot_dat <- dat[J,]
    return(metric_fun(boot_dat))
  }
  boot.out <- boot(data=dat, statistic=boot_metric_fun, R=B)
  confint <- boot.ci(boot.out, conf = conf.level, type = c('perc'))
  CI <- confint$percent[c(4,5)]
  return(CI)
}
set.seed(1)
boot_CI <- boot_CI_fun(dat=data, metric_fun=log_reg_fun)
boot_CI
```