---
title: "Solutions for chapter 2 exercises"
output: html_notebook
---

# Set up

```{r}
library(tidyverse)
library(psych)
library(rstudioapi) #To load data from local folder

### Setting the working directory to the parent folder of this script (Rstudio only)
sourceDir <- rstudioapi::getActiveDocumentContext()$path %>% str_extract("^.+/")
setwd(sourceDir)

set.seed(1)
options(scipen=10)

dat <- read_csv("AirCnC_MnM_exercises_data.csv")
```
# 0. The problem as seen by the PM
This preliminary section shows the calculations that the PM ran, for reference.
```{r}
# The booking rate is lower for customers who have seen the ad
dat %>% group_by(ad) %>% summarise(
  bkg_rate = mean(bkd)
)
```
0.b. What is the booking rate for customers who have seen the ad, restricting to customers considering an M&M property? Customers who haven’t seen the ad, with the same restriction?
```{r}
# This remains true even when restricting to customers considering an M&M property
dat %>% filter(mm==1) %>% group_by(ad) %>% summarise(
  bkg_rate = mean(bkd)
)
```

# 1. Understanding the behaviors

1.a. What are the behavioral categories for the variables in the data (Income, Ad, MM, Bkd)?

Income is a personal characteristic.
Ad is a business behavior.
MM is a customer behavior. 
Bkd is a customer behavior.

1.b. What is (are) the goal(s) of the ad?

The goals of the ad are 
- to increase the percentage of customers who consider an M&M property
- to increase the percentage of customers who book an M&M property

```{r}
# The ad increases the probability that a customer will consider an M&M property
mod_mm <- glm(mm~ad, family = 'binomial', data=dat)
summary(mod_mm)
```

```{r}
#The ad increases the probability that a customer will book an M&M property
dat <- dat %>%
  mutate(bkd_mm = bkd * mm) # Equal to 1 if and only if a customer books an M&M property
mod_bkd_mm <- glm(bkd_mm~ad, family = 'binomial', data=dat)
summary(mod_bkd_mm)
```
# 2. Resolving the mystery 
2.a. How does income affect these behaviors?
```{r}
# Income increases the probability that a customer will consider an M&M property
mod_mm <- glm(mm~income+ad, family = 'binomial', data=dat)
summary(mod_mm)
```

```{r}
# Income increases the probability that a customer will book an M&M property
mod_bkd_mm <- glm(bkd_mm~income+ad, family = 'binomial', data=dat)
summary(mod_bkd_mm)
```
2.b. What is the average income of customers considering an M&M property after seeing the ad? Without seeing the ad? 
```{r}
# Customers considering an M&M property after seeing the ad have a lower income than customers considering an M&M property without having seen the ad
dat %>% group_by(ad, mm) %>% summarise(
  avg_income = mean(income)
)
```
2.c. Can you now explain to the PM what happened?

The ad was effective at driving more customers to consider an M&M property across the board (i.e. irrespective of income). However, because there are more customers with a lower income than with a higher income, this added proportionately more lower-income customers to the pool of customers considering an M&M property. These lower-income customers have a lower likelihood to book a property, so the average booking rate across customers considering an M&M property decreased. In other words, the mix of customers considering an M&M property changed, but the individual probability that a customer would consider and book an M&M property increased. This ad is a resounding success!